Reviving stochasticity: SMBH-binary eccentricity is inherently random

### 2023-06-21 papers/paper-eccentricity/paper-eccentricity.tex
Section 2 checked

We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{rantala2017, rantala2020,mannerkoski2023}, implemented in \gadget{} \citep{springel2021}.
For details about the implementation of \ketju{}, we point the reader to \citet{mannerkoski2023}.
We use the same code configuration (including post-Newtonian corrections to order 3.5, fast multiple method with second order multipoles, and hierarchical time integration) as \citet{mannerkoski2023}.

We construct our galaxy models to represent the nuclear bulge of a gas-devoid elliptical galaxy, matching the models of \citet{nasim2020}.
Each galaxy is thus represented as a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
\begin{equation}\label{eq:dehnen}
    \rho_\star(r) = \frac{(3-\gamma)M_\star}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
\end{equation}
The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed with zero initial velocity.
We test five different mass resolutions by sampling the stellar phase space distribution with a varying number of stellar particles: $N_\star = \{2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$.

We then construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
The chosen eccentricities $e_0$ are 0.90 and 0.99.
The radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022} for the $e_0=0.90$ orbit.
For each orbital configuration, we run ten realisations for each mass resolution, to account for stochasticity caused by the discretised phase space.

>>>>

We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{mannerkoski2023,rantala2017}, in which the dynamics of
SMBH and stars in a small region around them are integrated with an algorithmically regularised integrator \citep{rantala2020}, whereas the dynamics of the remaining particles
is computed with the \gadget{} \citep{springel2021} fast multiple method (FMM) with second order multipoles. Together with hierarchical time integration this allows for
symmetric interactions and manifest momentum conservation. KETJU also includes Post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{Mora2004}.

Our galaxy models represent the nuclear bulge of a gas-devoid elliptical galaxy, matching the models of \citet{nasim2020}.
Each galaxy is thus represented as a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
\begin{equation}\label{eq:dehnen}
    \rho_\star(r) = \frac{(3-\gamma)M_\star}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
\end{equation}
The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed.
We test six different mass resolutions %by sampling the stellar phase space distribution
with a varying number of stellar particles: $N_\star = \{1.0\times10^5,2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$.

We then construct isolated merger initial conditions by placing two galaxies on two different elliptical orbits with eccentricities of $e_0=0.90$ and $e_0=0.99$, at a fixed
initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
The radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022} for the $e_0=0.90$ orbit.
For each orbital configuration, we run ten realisations for each mass resolution, to account for stochasticity caused by the discretised phase space.
